Polynomial expansions of finite volume data in a cut cell context

### 2023-04-18 example/DomainDecomposition/EBCellMerge/hoeb_geometry_test/_doc/paper.tex
initial ebcm framework compiles
added eigen comment and started documenting first example

Given a sufficiently smooth function $\psi$, we can approximate $\psi$
in the neighborhood of $\xbar$ using a Taylor expansion to order $P$:
\begin{equation}
\psi(\xbold)  =  \sum\limits_{p < P} C^p (\xbar - \xbar)^p \label{eqn::taylor}
\end{equation}
where $C^p$ this appropriate Taylor coefficient.  In three dimensions,
\begin{equation}
  C^p =\frac{1}{p!}
      \frac{\partial^{p_0}}{\partial x_0}
      \frac{\partial^{p_1}}{\partial x_1}
      \frac{\partial^{p_2}}{\partial x_2}  (\psi).
\end{equation}  
In finite volume  methods, we define grid data to be
averages over volumes.   So if we discretize the smooth function
$\psi$, the average over the volume $\ibold$ is given by
\begin{equation*}
 V_\ibold <\psi>_\ibold = \int\limits_{V_\ibold} \psi(\xbold) dV
\end{equation*}
If we insert the Taylor expansion \ref{eqn::taylor}, we get a discrete
approximation to the smooth function $\psi$ that is accurate to order $P$.
\begin{equation*}
 V_\ibold <\psi>_\ibold = \sum\limits_{p < P} C^p m^p. 
\end{equation*}
This forms a local polynomial expansion of $\psi$ around the volume
$\ibold$ expressed in terms moments ($m$) which are the natural
products of grid generation.  Modern, higher order embedded boundary
methods use this description directly to generate stencils
\cite{Overton2022a, Devendran2017, Schwartz2015}.  At the
core, all these algorithms have a similar procedure for finding the
Taylor coefficients $C^p$ which is worth discussing

@@

Consider a finite volume grid where we know all geometric moments to
order $P$.  Given values of $<\psi>$ everwhere on the grid, we wish to
approximate the Taylor coefficients at a volume $\ibold$ given that we
know $\cn_\ibold = \{\jbold\}$, the set of $N_v$ closest neighbors of
the volume $\ibold$.  Using the Taylor expansion at each point gives
us $N_v$ equations for $C^p$
\begin{equation}
  V_\jbold <\psi>_\jbold = \sum\limits_{p < P} C^p_\ibold m^p_\jbold.
\end{equation}
Let's say there are $N_p$ Taylor coefficients.  We can force $N_v >
N_p$ by getting a large enough neighborhood.  This means that we have
an overdetermined system of the form
\begin{equation*}
M C = P 
\end{equation*}
where $M = \{\{m^{pi}_\jbold\}\}$, $C=\{C^p\}$,  and $P= \{<\psi_\jbold>\}$.
Since the system is overdetermined, we can introduce a 
meaningful weighting matrix $W$ \footnotemark[3].  Choice of $W$
is crucial to the stability of these algorithms
\footnotemark[4].
We multiply and solve
\begin{equation*}
WMC = WP.
\end{equation*}
Since these matricies are (deliberately) not square, we approximate
the solution using a Moore-Penrose pseudo-inverse:
\begin{equation*}
C = ((WM)^T(WM))^{-1} (WM)^T P.
\end{equation*}
The matrix inversion at the heart of this method requires that we need
know to what degree the matrix $A \equiv (WM)^T(WM)$ is
well-conditioned.  Experience suggests that small cells play a role in
that conditioning and that the weighting matrix $W$ can be used to
mitigate conditioning issues that unweighted least squares may
produce.

@@

\footnotetext[7] {.}

\renewcommand{\thefootnote}{\fnsymbol{footnote}}\
\bibliography{references}

>>>>

Given a sufficiently smooth function $\psi$, we can approximate $\psi$
in the neighborhood of $\xbar$ using a Taylor expansion to order $P$:
\begin{equation}
\psi(\xbold)  =  \sum\limits_{p < P} C^p (\xbar - \xbar)^p \label{eqn::taylor}
\end{equation}
where $C^p$ this appropriate Taylor coefficient.  In three dimensions,
\begin{equation}
  C^p =\frac{1}{p!}
      \frac{\partial^{p_0}}{\partial x_0}
      \frac{\partial^{p_1}}{\partial x_1}
      \frac{\partial^{p_2}}{\partial x_2}  (\psi).
\end{equation}  
In finite volume  methods, we define grid data to be
averages over volumes.   So if we discretize the smooth function
$\psi$, the average over the volume $\ibold$ is given by
\begin{equation*}
 V_\ibold <\psi>_\ibold = \int\limits_{V_\ibold} \psi(\xbold) dV
\end{equation*}
If we insert the Taylor expansion \ref{eqn::taylor}, we get a discrete
approximation to the smooth function $\psi$ that is accurate to order $P$.
\begin{equation*}
 V_\ibold <\psi>_\ibold = \sum\limits_{p < P} C^p m^p. 
\end{equation*}
This forms a local polynomial expansion of $\psi$ around the volume
$\ibold$ expressed in terms moments ($m$) which are the natural
products of grid generation.  Modern, higher order embedded boundary
methods use this description directly to generate stencils
\cite{Overton2022a, Devendran2017, Schwartz2015}.  At the
core, all these algorithms have a similar procedure for finding the
Taylor coefficients $C^p$ which is worth discussing.

@@

Consider a finite volume grid where we know all geometric moments to
order $P$.  Given values of $<\psi>$ everwhere on the grid, we wish to
approximate the Taylor coefficients at a volume $\ibold$ given that we
know $\cn_\ibold = \{\jbold\}$, the set of $N_v$ closest neighbors of
the volume $\ibold$.  Using the Taylor expansion at each point gives
us $N_v$ equations for $C^p$
\begin{equation}
  V_\jbold <\psi>_\jbold = \sum\limits_{p < P} C^p_\ibold m^p_\jbold.
\end{equation}
Let's say there are $N_p$ Taylor coefficients.  We can force $N_v >
N_p$ by getting a large enough neighborhood.  This means that we have
an overdetermined system of the form
\begin{equation*}
M C = P 
\end{equation*}
where $M = \{\{m^{pi}_\jbold\}\}$, $C=\{C^p\}$,  and $P= \{<\psi_\jbold>\}$.
Since the system is overdetermined, we can introduce a 
meaningful weighting matrix $W$ \footnotemark[3].  Choice of $W$
is crucial to the stability of these algorithms
\footnotemark[4].
We multiply and solve
\begin{equation*}
WMC = WP.
\end{equation*}
Since these matricies are (deliberately) not square, we approximate
the solution using a Moore-Penrose pseudo-inverse:
\begin{equation*}
C = ((WM)^T(WM))^{-1} (WM)^T P.
\end{equation*}
This formulation can woven into the computation for stencil
components;  Devendran \cite{Dendran2017}, for example, explains the
process for Poisson's equation.

Unfortunately, the matrix $A \equiv (WM)^T(WM)$ can be poorly conditioned.
That these algorithms invert a matrix to compute
stencil coefficients introduces an interesting type of uncertainty: error in
the very coefficients of a stencil operation.   If the matrix is poorly
conditioned, the coefficients of the stencil can only be known to
limited precision.   This lack of precision is measurable as are its
consequences.

\section{All Regular geometry}

To investigate the consequences of this geometry, we consider as a
motivating example a geometry wherein all cells are uncut.

The matrix inversion at the heart of this method requires that we need
know to what degree the matrix $A \equiv (WM)^T(WM)$ is
well-conditioned.  Experience suggests that small cells play a role in
that conditioning and that the weighting matrix $W$ can be used to
mitigate conditioning issues that unweighted least squares may
produce.

@@

\footnotetext[7] {All matrix computations here are performed using
  Eigen \cite{eigenweb}, a beautifully designed and implemented
  package for on-processor linear algebra that is freely available.}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}\
\bibliography{references}
